### 2019-11-28 README.Rmd
some

---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 120)
```

# gosset

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/gosset)](https://cran.r-project.org/package=gosset)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
[![Build Status](https://travis-ci.org/agrobioinfoservices/gosset.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/gosset)
[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/master.svg)](https://codecov.io/github/agrobioinfoservices/gosset?branch=master)
[![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
<!-- badges: end -->

# *gosset*: Modelling Metadata and Experimental Citizen Science <img align="right" src="man/figures/logo.png">

## Overview

The **gosset** package provides an environment for modelling metadata and experimental citizen science data.

## Installation

The development version can be installed via

    library("devtools")
    devtools::install_github("agrobioinfoservices/gosset", upgrade = "never")

**gosset** is under development. Use with caution.

## Going further

The full functionality of **gosset** is illustrated in the package vignette. The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/) or from within R once the package has been installed.

## Contribution

You are welcome to contribute to this project. Please read our [contribution guideline](CONTRIBUTING.md).

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.