### 2020-01-08 README.Rmd
lifecycle-badge

---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 120)
```

# gosset

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/gosset)](https://cran.r-project.org/package=gosset)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
[![Build Status](https://travis-ci.org/agrobioinfoservices/gosset.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/gosset)
[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/master.svg)](https://codecov.io/github/agrobioinfoservices/gosset?branch=master)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

# *gosset*: Modelling Metadata and Experimental Citizen Science <img align="right" src="man/figures/logo.png">

## Overview

The **gosset** package provides the toolkit for a workflow on modelling metadata and experimental citizen science data, from collating data of different sources, model sellection and visualisation. The package is named after W.S. Gosset aka 'Student', a pioneer of modern statistics in small sample experimental design and analysis.

## Installation

The development version can be installed via

```r
library("devtools")
devtools::install_github("agrobioinfoservices/gosset", build_vignettes = TRUE)
```

## Example

## Going further

The full functionality of **gosset** is illustrated in the package vignette. The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/) or from within `R` once the package has been installed, e.g. via

```r
vignette("Overview", package = "gosset")
```

## Meta

* Please [report any issues or bugs](https://github.com/agrobioinfoservices/gosset/issues).

* License: GPL-3

* Get citation information for *gosset* in R by typing
`citation(package = "gosset")`.

* You are welcome to contribute to the *gosset* project. Please read our [contribution guide lines](CONTRIBUTING.md).

* Please note that the *gosset* project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By participating in the *gosset* project you agree to abide by its terms.